---
title: "Week 9 Lab - KEY"
author: "Prof. Amanda D. Perez, PhD"
output: rmdformats::material
---

# Simple & Multiple Regression: Mario Kart Data

```{r, warning = F, message = F}
library(tidyverse)
```

For this section you will be using Mario Kart Record data from TidyTuesday. See <https://github.com/rfordatascience/tidytuesday/tree/master/data/2021/2021-05-25> for more detailed information about the dataset.

```{r, warning = F, message = F}
mario <- read_csv("data/records.csv") 

glimpse(mario)
```

## Exercise 1 - 

In this first exercise we will be running a simple regression to predict `time` from `track`. As you can see below, there are `13` different tracks. Unlike with t-tests, we do NOT have to limit it to two levels. However, we do need to think about what our reference group will be. Generally, the reference group is either a baseline (because all other levels will be compared to it) or it is the level you are most interested in. 



```{r}
mario$track <- as.factor(mario$track)
levels(mario$track)
```

Right now, `Banshee Boardwalk` is the default reference level. We know this because it is the first level that appears above. I want to change the reference level to `Bowser's Castle`. I will do that in the below code.

```{r}
mario$track <- relevel(mario$track, ref = "Bowser's Castle")
```

Below, I will confirm that `Bowser's Castle` is the new reference level.

```{r}
levels(mario$track)
```


`Bowser's Castle` shows up first and is now our reference level. We can now move on to our simple regression. Fill in the formula below to run a regression predicting `time` (the DV) from `track` (the IV). When you have correctly specified the `lm()` model, the `summary(mod.mario.simple)` will run without error.

```{r}
mod.mario.simple <- lm(time ~ track, data = mario)
summary(mod.mario.simple)
```

Your job is to interpret:

**The intercept**
[RESPONSE HERE]


**All significant slopes**
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]


**The R-Squared**
[RESPONSE HERE]

## Exercise 2 - 

Next we will visualize the regression model. I have provided a basic graph for your below. Your job is to update the graph and: fill each bar by track, add a theme, hide the legend using `guides` argument, label the x and y axes, and add a title. Interpret what you see.

```{r}
ggplot(mario, aes(track, time, fill = track)) +
  stat_summary(fun = "mean", geom = "bar") +
  coord_flip() +
  labs(x = "Track", y = "Time in seconds", title = "Mario Kart Track Times") +
  guides(fill = "none") +
  theme_minimal()
```

## Exercise 3 - 

Now that we have run a simple regression (regression with just one IV) we want to now run a multiple regression. We want to run a new model that will add on a second IV: `type`

```{r}
mario$type <- as.factor(mario$type)
levels(mario$type)
```

In this case, the reference level is `Single Lap`. I like this and will not change the reference level.

Now, it is time to build our multiple regression model. We will be predicting `time` (DV) from `track` (IV 1) AND `type` (IV 2). When you have correctly specified the `lm()` model, the `summary(mod.mario.multiple)` will run without error.

```{r}
mod.mario.multiple <- lm(time ~ track + type, data = mario)
summary(mod.mario.multiple)
```

Your job is to interpret:

**The intercept**
[RESPONSE HERE]


**All significant slopes**
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]

**The R-Squared**
[RESPONSE HERE]

## Exercise 4 - 

Next we will visualize the multiple regression model. I have provided a basic graph for your below. Your job is to update the graph and: fill each bar by type, add a theme, and label the x and y axes, label the fill title (legend), and add a title. Interpret what you see.

```{r}
ggplot(mario, aes(track, time, fill = type) )+
  stat_summary(fun = "mean", geom = "bar", position = "dodge") +
  coord_flip() +
  labs(x = "Track", y = "Time in seconds", fill = "Type", title = "Mario Kart Track Times")  +
  theme_minimal()
```


# Simple & Multiple Regression: Fishing Data

```{r, warning = F, message = F}
library(tidyverse)
```

For this section you will be using Great Lakes Fishery Commission data from TidyTuesday. See <https://github.com/rfordatascience/tidytuesday/tree/master/data/2021/2021-06-08> for more detailed information about the dataset.

```{r, warning = F, message = F}
set.seed(5678)
fishing <- read_csv("data/fishing.csv") %>%
  sample_n(8000)

glimpse(fishing)
```

## Exercise 5 - 

We will be running a simple regression to predict `grand_total` from `species`. As you can see below, there are `51` different species. Unlike the previous simple regression, I am only interested in 5 different fish species: Buffalo, Crappie, Herring, White Bass, and Yellow Perch. 


```{r}
fishing$species <- as.factor(fishing$species)
levels(fishing$species)
```

Below I am filtering the data to only include these levels:

```{r}
fishing <- fishing %>%
  filter(species == "Buffalo" | species == "Crappie" | species == "Herring" | species == "White Bass" | species == "Yellow Perch")

fishing$species <- droplevels(fishing$species)
levels(fishing$species)
```


Above we see that `Buffalo` is the reference level. However, I want `White Bass` to be the reference level:

```{r}
fishing$species <- relevel(fishing$species, ref = "White Bass")
```

Below, I will confirm that `White Bass` is the new reference level.

```{r}
levels(fishing$species)
```


`White Bass` shows up first and is now our reference level. We can now move on to our simple regression. Fill in the formula below to run a regression predicting `grand_total` (the DV) from `species` (the IV). When you have correctly specified the `lm()` model, the `summary(mod.fish.simple)` will run without error.

```{r}
mod.fish.simple <- lm(grand_total ~ species, data = fishing)
summary(mod.fish.simple)
```

Your job is to interpret:

**The intercept**
[RESPONSE HERE]


**All significant slopes**
[RESPONSE HERE]
[RESPONSE HERE]


**The R-Squared**
[RESPONSE HERE]

## Exercise 6 - 

Next we will visualize the regression model. This time, you must write your own ggplot code. Make sure you: fill each bar by species, add a theme, hide the legend using `guides` argument, label the x and y axes, and add a title. Interpret what you see.

```{r, warning = F, message = F}
ggplot(fishing, aes(species, grand_total, fill = species)) +
  stat_summary(fun = "mean", geom = "bar")  +
  labs(x = "Fish Species", y = "Mean Grand Total", title = "Grand Total Fish Species") +
  guides(fill = "none") +
  theme_minimal()
```

## Exercise 7 - 

Now that we have run a simple regression (regression with just one IV) we want to now run a multiple regression. We want to run a new model that will add on a second IV: `lake`

```{r}
fishing$lake <- as.factor(fishing$lake)
levels(fishing$lake)
```

In this case, the reference level is `Erie`. I like this and will not change the reference level.

Now, it is time to build our multiple regression model. We will be predicting `grand_total` (DV) from `species` (IV 1) AND `lake` (IV 2). When you have correctly specified the `lm()` model, the `summary(mod.fish.multiple)` will run without error. This is running an INTERACTION Model. You can ignore the rows with `NA`s. The NA rows mean there are not data points for that specific combination of data points.

```{r}
mod.fish.multiple <- lm(grand_total ~ species * lake, data = fishing)
summary(mod.fish.multiple)
```

Your job is to interpret:

**The intercept**
[RESPONSE HERE]


**All significant slopes (main effects only)**
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]
[RESPONSE HERE]

**The R-Squared**
[RESPONSE HERE]

**How many significant interactions are there?**


## Exercise 8 - 

Next we will visualize the multiple regression model. This time, you must write your own ggplot code. Make sure you: make species the x axis, fill each bar by lake, add a theme,  label the x and y axes, label the fill title (legend), and add a title. Interpret what you see.

```{r, warning = F, message = F}
ggplot(fishing, aes(species, grand_total, fill = lake) )+
  stat_summary(fun = "mean", geom = "bar", position = "dodge") +
  labs(x = "Fish Species", y = "Mean Grand Total", fill = "Lake", title = "Fish Species")  +
  theme_minimal()
```