---
title: "Week 11 Lab"
author: "Prof. Amanda D. Perez, PhD"
output: rmdformats::material
---

# Cross Validation: The Office Data

```{r, warning = F, message = F}
library(tidyverse)
library(caret)
```

For this section you will be using data about the U.S. version of the television show *The Office*. 

```{r, warning = F, message = F}
office_df <- read_csv("data/the_office.csv") 
```


## Exercise 1: Partitioning the data.
Create an index of for our data partitioning. 
We are doing an 80/20 split so set `p = .8`.
We do not want a list (we want a matrix) so set `list = FALSE`.
Set `times = 1`
Fill in the blank such that you are looking at `office_df$imdb_rating` in the first argument.


```{r, error = TRUE}
set.seed(1234) #set the seed so we can all reproduce results
index <- createDataPartition(office_df$____, )
```



## Exercise 2: Create our testing & training datasets train_df and test_df

Below we are creating `office_train` (holds our data to use to train our cv model) and `office_test` (what we will use to test our final model). Correctly fill in the blanks using the index object we created. 

```{r train-test, error = TRUE}
office_train <- office_df[___, ]
office_test  <- office_df[-___, ]
```

## Exercise 3: Specify the type of training method/numbers of folds to specify

Now it is time for us to specify our cross-validation. Fill in the blanks below and:

Use the `trainControl` function.
Specify "cv" as the method.
Set `number = 5`.
Set `savePredictions = "all"`.

```{r, error = TRUE}
ctrl_specs <- ____(method = "___", number = ___, 
                           savePredictions = "___")
```



## Exercise 4: Run our linear model

Fill in the blanks for us to run our linear model via cross-validation. 
Fill in the blanks and:
Specify `imdb_rating` as the DV
Specify a main effect only model (using '+' signs) for the following IVs: season, episode, michael
Set `data = office_train`.
Set `method = "lm"`
Set trControl = ctrl_specs

```{r, error = TRUE}
#set another random seed for our random assignment in our folds
set.seed(1990)

model.1 <- train([DV] ~ [IV1] + [IV2] + [IV3],
                 data = ___, method = "___", trControl = ___)
```



## Exercise 5: Model Outputs 

Below, run the code to print the summary for model.1. Interpret the RSquared value.

```{r, error = TRUE}
print(model.1)
```


Below, run the code to print the statistics for each fold. We ran 5 folds. Interpret the RSquared value among the folds. Which fold performed the best and which fold performed the worst?

```{r, error = TRUE}
model.1$resample
```

## Exercise 6: See output in terms of regression coefficients

Interpret the linear regression output (intercept, all slopes, adjusted RSquared)

Reminder: `michael` is a binary variable (being treated as a factor). The 0 = michael scott is not in the episode and the 1 = michael scott is in the episode. 

```{r fit,  eval=FALSE}
summary(model.1)
```

ANSWER:

## Exercise 7: Assess variable importance of the predictors

Run the below code. Which variable is most important and which variable is least important?

```{r, error = TRUE}
varImp(model.1) 
```

ANSWER:

## Exercise 8: Apply our model to the office_test dataframe

Fill in the blank in order to predict the outcome using the model from `office_train` applied to the `office_test` data.

```{r, error= TRUE}
office_predictions <- predict(model.1, newdata = ___) 
```

## Exercise 9: Evaluate Model Metrics

Run the below code and interpret the RSquared and RMSE.

```{r, error = TRUE}
data.frame(R2 = R2(office_predictions, office_test$imdb_rating),
            RMSE = RMSE(office_predictions, office_test$imdb_rating))

```

ANSWER:

## Exercise 9: Compare our Training Metrics to our Testing Matrix

The below code will help you compare the r squared and RMSE from our training versus testing model. Interpret the output below.

```{r, error = TRUE}
data.frame(model.1$results[3],
            R2_testing = R2(office_predictions, office_test$imdb_rating),
            model.1$results[2],
            RMSE_testing = RMSE(office_predictions, office_test$imdb_rating))

```

ANSWER: