---
title: "Finding & Reading the Research"
subtitle: "<br><br> Psych 101x"
author: "Amanda D. Perez, PhD"
output:
  xaringan::moon_reader:
    css: ["https://amandadperez.github.io/TutoRials/docs/Personal/BASIS/xaringan-themer.css", "https://amandadperez.github.io/TutoRials/docs/Personal/BASIS/slides.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      countIncrementalSlides: false
---


```{r child = "setup.Rmd"}
```


```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(emo)
xaringanExtra::use_scribble()
# highlightStyle: solarized-dark

library(viridis)
library(sugrrants)
library(lubridate)
```

## Finding Scientific Sources

--

PsycINFO

--

**UC Berkeley Library Guides**

https://guides.lib.berkeley.edu/ 

---

## Finding Scientific Sources

**TIPS:**

--

Use *“AND”* *“OR”*

  Example: Depression *AND* adolescent *OR* teenager
  
--

Use '*' for all possible word endings

  Example:   
  Depres*
  
  adolescent*
  
  teen*


---

## Reading the Research

**Components of an empirical journal article**

---

## Abstract

Summary of the article; read it to figure out what the article is about

---

## Introduction

Explains the topic of the study, describes the theoretical and empirical background for the research, and states the specific research goals for the current study 

---

## Method

How the researchers conducted the study; includes subsections on participants, materials, and procedure

---

## Results

Describes the statistical tests used and the quantitative or qualitative results of the study

---

## Discussion

Summarizes the study and discusses the study’s significant contributions and perhaps alternative explanations for their results

---

## References

Full bibliographic listing of all the sources the authors cited in the article 

---

## Reading with a Purpose

**What is the argument?** 

First read the abstract, then skip to the end of the introduction to read about the goals of the current study, then read the rest of the introduction. You might also read the first paragraph of the discussion section.

--

**What is the evidence to support the argument?**

Read the method and results sections.

---

## TOOLS FOR CONSUMING RESEARCH

---

## Variables

A **variable** is something that changes or varies, so it needs to have at least two levels or values (but it can have more). 

--

**Example:** "80% of college undergraduates have taken a selfie." 

Selfie is the variable, and there are **two levels**: those who have taken a photo of themselves electronically, and those who have not. 

--

**Example:** “People who exercise regularly are happier than others.”  

There are **two variables** here: **exercising regularly** (exercise at least once a week or do not exercise at least once a week) and **happiness** (some numerical score on a happiness inventory).


---

## Variable

**Next, we will look at different types of variables:** 

Measured variables

Manipulated variables

Conceptual variables

Operationalized variables

---

## Measured VS Manipulated Variable

**Measured variable:** Levels are observed and recorded by the researcher. 

**Examples** of measured variables include height, IQ, gender, and hair color. Psychologists also measure more abstract variables such as depression and stress.

---

## Measured VS Manipulated Variable

**Manipulated variable:** The researcher controls a variable, usually by assigning participants to different levels of that variable. 

**Example** of a manipulated variable: a researcher might assign some people to take a test in a room with many other people and assign others to take the test alone. 

---

## Measured VS Manipulated Variable

Some variables **can only be measured**—not manipulated.

You cannot assign participants to be a particular age, as age is a      naturally occurring variable. 

Also, sometimes it is unethical to manipulate variables. For ex, in a study on the long-term effects of elementary education, you could not ethically assign children to “high-quality school” and “low-quality school” conditions. 

---

## Measured VS Manipulated Variable

Some variables can be **either** manipulated or measured.  

For example, if childhood extracurricular activities were the variable of interest, you could measure whether children already take music or drama lessons, *or* you could manipulate this variable by assigning some children to take music lessons and others to take drama lessons. 

---

## Operational Definition

In order to test hypotheses with empirical data, researchers need to develop **operational definitions**, or operational variables. 

To operationalize is to **turn a conceptual definition into a measured or manipulated variable**. 

This is *how* you will measure your variable

--

**Example:** spending time socializing is a conceptual variable, and how often a person spends an evening alone, socializes with friends, and sees relatives in a typical week are operational variables. 











